---
title: "L01 Introduction"
subtitle: "Data Visualization (STAT 302)" 
author: "Pablo Restrepo"
pagetitle: "L01 Pablo Restrepo"
format:
  html:
    toc: true
    toc-location: left
    embed-resources: true
    link-external-newwindow: true
execute:
  warning: false
from: markdown+emoji
reference-location: margin
citation-location: margin
editor_options: 
  chunk_output_type: console
---

::: {.callout-tip icon="false"}
## Github Repo Link

<https://github.com/pablorestrepo2027-art/STAT-302.git>
:::

## Goals

1.  Ensure software for this course is properly installed and functional
2.  Understand the importance of visualization
3.  Develop and follow a proper workflow
4.  Construct and explore a few plots using `ggplot2`
5.  Add layers to a plot that include title, axis, and pre-set themes

**Don't worry if you cannot do everything here by yourself.** You are just getting started and the learning curve is steep, but remember that the instructional team and your classmates will be there to provide support. Persevere and put forth an honest effort and this course will payoff.

```{r}
#| label: load-packages

# Load package(s) 
library(tidyverse)
library(skimr)
library(devtools)

# Data can go here as well
load("data/legosets.rda")
```

::: callout-tip
## Organization is important!

1.  All packages should always be loaded at the TOP of the document.

2.  All data should be in your `/data` subdirectory. Data loading should go at the top of the document under the libraries.
:::

## Tasks

Complete the following tasks and **include some indication that you have completed each** in your document.

### **Task 1: Download R :white_check_mark:**

### **Task 2: Download RStudio :white_check_mark:**

### **Task 3: Install Packages :white_check_mark:**

### **Task 4: Create GitHub and Setup Git:white_check_mark:**

### **Task 5: Create L00 :white_check_mark:**

### **Task 6: Explore Quarto and ggplot2 :white_check_mark:**

## Exercise 1

Let's delve into why data visualizations are important.

### Visualization 1

Read the following article and answer the following questions.

[bars_vs_grocery_stores](https://flowingdata.com/2014/05/29/bars-versus-grocery-stores-around-the-world/)

1)  List one thing you learned from the article?

::: {.callout-tip icon="false"}
## Solution

I learned that North Dakota and Montana are #1 and #2 in bars per capita in the United States. This surprised me, however it makes sense given that these states have small populations compared to the rest of the US.
:::

2)  How did the main visualization (United States) enhance the article?

::: {.callout-tip icon="false"}
## Solution

It added more nuance to the original proposition of bars vs grocery stores by adding a gradient. It added more information by visualizing the ratios in each area of bars to grocery stores.
:::

3)  List at least one thing you liked about the \[United States\] graphic and one thing you think could be improved.

::: {.callout-tip icon="false"}
## Solution

I liked the color scheme, it made the gradient very clear and intuitive. Especially with the implied associations (more brown/orange for bars, similar to wood and beer. Green for grocery stores with vegetables). One thing that can be improved the US map itself. Adding state lines would make it easier to tell what states have more bars or grocery stores.
:::

### Visualization 2

Open this [dashboard](https://edgarruiz.shinyapps.io/db-dashboard/) and answer the following questions.

1)  What story is this dashboard telling and how do the visualizations enhance that?

::: {.callout-tip icon="false"}
## Solution

The dashboard recaps annual flight statistics for Airlines that operate in the United States. The visualizations enhance this recap by facilitating comparisons between stats like destination airport and months within airlines and also between airlines through stats like flights per day on an yearly basis or within a specific month.
:::

2)  What might a dashboard provide that a static graphic cannot.

::: {.callout-tip icon="false"}
## Solution

A dashboard facilitates comparisons thanks to its interactivity, which enables users to filter and search for the data they are looking for. Dashboards also provide multiple perspectives/data representations. In this case, the Flights Dashboard provides both bar graphs and tables with notable statistics.
:::

## Exercise 2

Let's look at some interesting patterns in the history of LEGO! We'll be using data taken from the [`lego` package](https://github.com/seankross/lego). A copy of the data is located here `data/legosets.rda`.

First, let's look at the average cost of LEGO sets over time. The main variable of interest here is `USD_MSRP`, or the manufacturer's suggested retail price in constant dollars. Let's recreate the following graphic.

::: {.callout-tip icon="false"}
## Solution

```{r}
yearly_price <- legosets %>% 
  group_by(Year) %>% 
  summarise(avg_usd_msrp = mean(USD_MSRP, na.rm=TRUE),
            med_usd_msrp = median(USD_MSRP, na.rm=TRUE),
            n = n() - sum(is.na(USD_MSRP))
            )


ggplot(
  data = yearly_price, 
  mapping = aes(x = Year, y = avg_usd_msrp)
  ) + 
  geom_line() + 
  geom_point() + 
  theme_minimal() +
  labs(
    title = "Average Annual Price of LEGO Sets", 
    subtitle = "Amounts are reported in US Dollars (USD)",
    y = "Price (USD)",
    caption = "Source: LEGO"
    )
```
:::

## Exercise 3

Let's explore how many unique themes LEGO now offers.

```{r}
#| label: ex3
#| echo: false

theme_counts <- legosets |>
  count(Theme)

theme_counts |>
  ggplot(aes(y = Theme, x = n)) +
  geom_col() +    #or geom_bar(stat = "identity") +
  labs(
    x = "Number of Sets",
    y = NULL
  ) +
  theme_minimal()
```

What is bad/ugly about the plot above?

::: {.callout-tip icon="false"}
## Solution

All the labels are overlapping, making them hard to read. It is difficult to identify the highest and lowest, there is no order to the plot. Furthermore, it is difficult to tell which bar belongs to which label because of the overlap and the size of the bars. The x-axis numbers also need to be more detailed as it is hard to tell the number of sets for each theme.
:::

Now remake the plot BUT with the improvements.

::: {.callout-tip icon="false"}
## Solution

```{r}
theme_counts %>%
  filter(n >= 100) %>% 
  arrange(n) %>% 
  mutate(Theme = fct_inorder(Theme)) %>% 
  ggplot(aes(y = Theme, x = n)) +
  geom_col() +
  labs(
    x = "Number of Sets",
    y = NULL
  ) +
  theme_minimal()


```

The Chart shows that the most popular themes (in terms of number of sets) are Duplo, Star Wars, and Town. Duplo is the highest by a healthy margin, being the only theme with over 400 sets. The least popular themes that are still above 100 sets are Explore, Ninjago, and Friends.
:::

## Exercise 4

Now you try, let's look at how the number of pieces per set has changed over time. Because Duplo sets are much smaller (since they're designed for toddlers), we made a special indicator variable for them in the code below.

Use an appropriate graph to plot the average number of pieces over time. Have either separate lines or separate plots for the Duplo and Lego categories. Be sure the title, x-axis, and y-axis are properly labeled.

```{r}
#| label: ex4-data-wrangle

pieces_per_year <- legosets %>%
  # new indicator variable
  mutate(Duplo = ifelse(
    Theme == "Duplo",
    "Duplo", 
    "Lego")) %>% 
  # group_by() both year and Duplo
  group_by(Year, Duplo) %>% 
  # summarize() mean pieces
  summarise(avg_pieces = mean(Pieces, na.rm=TRUE),
            med_pieces = median(Pieces, na.rm=TRUE),
            n = n() - sum(is.na(Pieces))
            ) %>% 
  ungroup()


pieces_per_year %>% 
  filter(Year > 1990) %>% 
ggplot(
  mapping = aes(x = Year, y = avg_pieces, color = Duplo)
  ) + 
  geom_line() + 
  geom_point() + 
  theme_minimal() +
  labs(
    title = "Average Number of Pieces in a LEGO Set Over Time", 
    y = "Average Pieces in a LEGO Set",
    caption = "Source: LEGO"
    )
```

What trends do you notice in the plots? Are you surprised by the plots?

::: {.callout-tip icon="false"}
## Solution

The plots show that both duplo and other LEGO sets have increased in average number of pieces from 1990 to 2015. Both experienced some drops in average pieces, Duplo especially in the early 2000s. On the other hand, both LEGO and Duplo sets had a significant increase in average pieces around 2005-2007. I am surprised by how basic Duplo sets were in the early 90s and by the spike in the average pieces in LEGO sets around the mid-2000s.
:::
